\Large\textbf{Surrogate Regret Bounds for Polyhedral Losses}

### 2021-12-03 rates-neurips-21/neurips-poster/main.tex
pass for wordiness

\documentclass[t,mathserif,11pt,aspectratio=1610,handout]{beamer}
% option "t" means align top
% option "handout" for handout mode (each frame takes only one page)
% option "notes" to add notes pages

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The short version}{}
  Surrogate risk minimization for supervised learning:
  \pause
  \begin{itemize}[<+->]
    \item Goal: optimize \redem{discrete ``target''} loss $\ell(r,y)$  \hintright{e.g. classification, structured predictions, \dots}  \\
          \hint{$r = $ prediction in a finite set, $y = $ label}
    \item Approach: optimize \blueem{continuous ``surrogate''} loss $L(u,y)$  \hintright{then ``link'' to target prediction $r = \psi(u)$} \\
          \hint{$u = $ prediction in $\reals^d$}
  \end{itemize}

  \vskip2em
  Question: when does (quick) \blueem{surrogate} convergence imply (quick) \redem{target} convergence?  \\
  \hint{``Surrogate regret bounds'' or ``regret transfer rates''}

@@

  \vskip2em
  This paper: \textbf{polyhedral surrogates}  \hintright{piecewise-linear and convex}
\end{frame}

\begin{frame}{~}{}
  \vskip-1em
  \centerline{\includegraphics[width=0.75\linewidth]{figs/hinge}}
  \vskip1em

  \textbf{Polyhedral:} In all cases, \blueem{surrogate regret} $\leq \epsilon$ $\implies$ \redem{target regret} $\leq O(\epsilon)$

  \smallskip
  \enumitem{\bf Theorem 1:} All polyhedral surrogates have \textbf{linear} regret transfer rates!

@@

  \textbf{Non-polyhedral:} There exist cases where \blueem{surrogate regret} $\leq \epsilon$ yet \redem{target regret} $\geq \Omega(\sqrt{\epsilon})$.

  \smallskip
  \enumitem{\bf Theorem 2:} All sufficiently ``non-polyhedral'' surrogate transfers are \textbf{quadratically slower}.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Supervised learning, surrogate risk minimization}{}
  \textbf{Data:} $(x,y) \in \X \times \Y$   \hintright{from distributions $\D$}

@@

  \vskip1em
  \pause
  \redem{Discrete target loss:} $\ell: \R \times \Y \to \reals_{\geq 0}$.  \hintright{discrete: $\R$ is finite}

  \only<4>{
    \vskip1em
    Examples:
    \begin{itemize}
      \item Classification ($0-1$ loss)     \hintright{$\R = \Y$}
      \item Classification with abstention  \hintright{$\R = \Y \cup \{\bot\}$}
      \item Ranking \hintright{$\R = $ permutations of $\Y$}
      \item Top-$k$ \hintright{$\R = $ size-$k$ subsets of $\Y$}
    \end{itemize}
  } % end only

@@

  \pause
  \vskip1em
  \blueem{Continuous surrogate loss:} $L: \reals^d \times \Y \to \reals_{\geq 0}$.  \hintright{for some $d$}

  \pause
  \vskip1em
  Examples:
  \begin{itemize}
    \item Hinge loss for classification ($0-1$ loss)
    \item BEP surrogate for classification with abstention  \hintright{Ramaswamy et al. 2018}
    \item ...
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{What makes for a good \blueem{surrogate}?}{}
  \pause
  \textbf{1.} Optimizable. \\
  \hint{Convex; strongly convex, etc.}

  \pause
  \vskip1em
  \textbf{2.} Generalization bound; convergence rate. \\
  \hint{e.g. with $n$ data points, $\Reg_L(h;\D) \leq O(\sqrt{1/n})$.}

  \pause
  \vskip1em
  \textbf{3. Connection to the \redem{target}.}

  \pause
  \vskip1em
  Formalize via \textbf{regret transfer function} (also called \emph{calibration function}) $\zeta: \reals_{\geq 0} \to \reals_{\geq 0}$:  \hintright{continuous at zero, $\zeta(0) = 0$}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Polyhedral losses}{}
  A function is \textbf{polyhedral} if it is the pointwise maximum of a finite set of affine functions.  \\
  \hint{``Piecewise-linear convex''}

  \pause
  \vskip1em
  The \blueem{surrogate loss} $L: \reals^d \times \Y \to \reals_{\geq 0}$ is \textbf{polyhedral} if each $L(\cdot ,y)$ is polyhedral.

  \pause
  \vskip1em
  Recent work: \blueem{polyhedral surrogates} are natural convexifications of \redem{discrete losses}.
  \begin{itemize}[<+->]
    \item Finocchiaro, \textbf{Frongillo}, \textbf{Waggoner} 2019:
    \begin{itemize}
      \item Every \redem{discrete loss} $\ell$ has a consistent \blueem{polyhedral surrogate} $(L,\psi)$.
      \item Further, that surrogate is obtained by ``embedding'' $\ell$ in $\reals^d$.
    \end{itemize}
    \item Finocchiaro, \textbf{Frongillo}, \textbf{Waggoner} 2020: investigates dimensionality $d$
    \item Inspiration: BEP surrogate of Ramaswamy et al. 2018
    \item Recent work, applications: Wang and Scott 2020 on Westin-Watkins hinge
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
  \begin{block}{Theorem 1}
    Suppose $L$ is polyhedral and $(L,\psi)$ are consistent for the discrete target $\ell$.
    Then there exists $C > 0$ such that, for all $\D$ and $h$,
    \[ \Reg_{\ell}(\psi \circ h; \D) \leq C \cdot \Reg_L(h; \D) . \]
  \end{block}

  \pause
  \vskip1em
  \emph{Proof sketch.}
  Fix $x$; let $p,q \in \simplex$ be distributions of $y$ given $x$.
  \begin{itemize}[<+->]
    \item Suffices to prove that for all $p$ and $u$, $\Reg_{\ell}(\psi(u);p) \leq C \cdot \Reg_L(u;p)$.
    \item Consistent $\implies$ for all $q$, there exists $\alpha_q > 0$ such that $\Reg_{\ell}(\psi(u); q) \leq \alpha_q \cdot \Reg_L(u; q)$.
    \item Polyhedral losses have a \textbf{finite structure}: \hintright{Closely related to embeddings framework}
    \begin{itemize}
      \item There is a finite $U \subseteq \reals^d$ such that, for all $p$, $U$ contains an optimal prediction.
      \item These $U$ partition $\simplex$ into finitely many polytopes.
    \end{itemize}
    \item Regret is linear on each polytope.
    \item Therefore $\Reg_L(u ; p)$ is a convex combination over the corners.
    \item So we can take $C = \max_{\text{corners $q$}} \alpha_q$.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
  \begin{block}{Theorem 2}
    Let $L$ be a locally strongly convex surrogate with locally Lipschitz gradient.
    Suppose $(L,\psi)$ is consistent for $\ell$ with:
    \[  \Reg_{\ell}(\psi \circ h) \leq \zeta\Big( \Reg_L(h) \Big) . \]
   Then there exists $c > 0$ such that, for all small enough $\epsilon > 0$,
    \[ \zeta(\epsilon) \geq c \sqrt{\epsilon} . \]
  \end{block}

  \pause
  \emph{Proof idea:}
  \begin{itemize}
    \item Fix a ``boundary'' prediction $u_0$.
    \item Consider a conditional distributions $\{p_{\lambda} : 0 \leq \lambda \leq 1\}$
    \item \redem{Target regret} shrinks linearly as $\lambda \to 0$
    \item But \blueem{Surrogate regret} shrinks with $\sqrt{\lambda}$.
  \end{itemize}

  \pause
  Example: exponential loss,  \hintright{not strongly convex, but locally s.c.}  \\
  Huber loss                  \hintright{via a strengthening of the theorem}
\end{frame}

>>>>

\documentclass[t,mathserif,12pt,aspectratio=1610,handout]{beamer}
% option "t" means align top
% option "handout" for handout mode (each frame takes only one page)
% option "notes" to add notes pages

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The short version}{}
  Surrogate risk minimization for supervised learning:
  \pause
  \begin{itemize}[<+->]
    \item Goal: optimize \redem{discrete ``target''} loss $\ell(r,y)$  \hintright{classification, structured prediction}  \\
          \hint{$r = $ prediction, $y = $ label}
    \item Approach: optimize \blueem{continuous ``surrogate''} loss $L(u,y)$  \hintright{then ``link'' to target prediction $r = \psi(u)$} \\
%          \hint{$u = $ prediction in $\reals^d$}
  \end{itemize}

  \vskip1em
  \textbf{Q:} When does (quick) \blueem{surrogate} convergence imply (quick) \redem{target} convergence?  \\
  \hint{``Surrogate regret bounds'' or ``regret transfer rates''}

@@

  \vskip1em
  This paper: \textbf{polyhedral surrogates}  \hintright{piecewise-linear and convex}
\end{frame}

\begin{frame}{~}{}
  \vskip-1em
  \centerline{\includegraphics[width=0.65\linewidth]{figs/hinge}}
  \vskip1em

  \enumitem{\bf Theorem 1:} All polyhedral surrogates have \textbf{linear} regret transfer rates!

  \smallskip
  \textbf{Polyhedral:} \blueem{surrogate regret} $\leq \epsilon$ $\implies$ \redem{target regret} $\leq O(\epsilon)$.

@@

  \enumitem{\bf Theorem 2:} Sufficiently ``non-polyhedral'' transfers are \textbf{quadratically slower}.

  \smallskip
  \textbf{Non-polyhedral:} Can have \blueem{surrogate regret} $\leq \epsilon$ yet \redem{target regret} $\geq \Omega(\sqrt{\epsilon})$.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Background: surrogate risk, polyhedral losses}{}
  \textbf{Data:} $(x,y) \in \X \times \Y$   \hintright{from distributions $\D$}

@@

  \vskip1em
  \pause
  \redem{Discrete target loss:} $\ell: \R \times \Y \to \reals_{\geq 0}$.  \hintright{discrete: $\R$ is finite}  \\
  \hint{Classification ($0-1$ loss), ranking, top-$k$}

@@

  \pause
  \vskip1em
  \blueem{Continuous surrogate loss:} $L: \reals^d \times \Y \to \reals_{\geq 0}$.  \hintright{for some $d$} \\
  \hint{Hinge loss for classification, BEP surrogate (Ramaswamy et al. 2018)}

  \vskip1em
  \blueem{Polyhedral surrogate} $L: \reals^d \times \Y \to \reals_{\geq 0}$: \\
  pointwise maximum of a finite set of affine functions.

  \pause
  \vskip1em
  Recent work: \blueem{polyhedral surrogates} are natural convexifications of \redem{discrete losses}.
  \begin{itemize}[<+->]
    \item Finocchiaro, Frongillo, Waggoner 2019, 2020; applications e.g. Wang, Scott 2020.
%    \begin{itemize}
%      \item Every \redem{discrete loss} $\ell$ has a consistent \blueem{polyhedral surrogate} $(L,\psi)$.
%      \item Further, that surrogate is obtained by ``embedding'' $\ell$ in $\reals^d$.
%    \end{itemize}
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{~}{}
  Desirable surrogates:
  \begin{itemize}
    \item Optimizable.  \hintright{Convex, strongly convex, etc.}
    \item Generalization bound, convergence rate.
%  \textbf{2.} Generalization bound; convergence rate. \\
%  \hint{e.g. with $n$ data points, $\Reg_L(h;\D) \leq O(\sqrt{1/n})$.}
    \item \textbf{Connection to \redem{target}...}
  \end{itemize}

  \vskip1em
  \textbf{Regret transfer function} (\emph{calibration function}) $\zeta: \reals_{\geq 0} \to \reals_{\geq 0}$:  \hintright{continuous at zero, $\zeta(0) = 0$}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Results}{}
  \begin{block}{Theorem 1}
    Suppose $L$ is polyhedral and $(L,\psi)$ are consistent for the discrete target $\ell$.
    Then there exists $C > 0$ such that, for all $\D$ and $h$,
    \[ \Reg_{\ell}(\psi \circ h; \D) \leq C \cdot \Reg_L(h; \D) . \]
  \end{block}

  \pause
  \vskip1em
  \emph{Proof sketch.}
  Fix $x$; let $p,q \in \simplex$ be distributions of $y$ given $x$.
  \begin{itemize}[<+->]
    \item Suffices to prove that for all $p$ and $u$, $\Reg_{\ell}(\psi(u);p) \leq C \cdot \Reg_L(u;p)$.
    \item $\forall q$, $\exists \alpha_q > 0$ such that $\Reg_{\ell}(\psi(u); q) \leq \alpha_q \cdot \Reg_L(u; q)$.
    \item Polyhedral losses have a \textbf{finite structure}: \hintright{related to embeddings framework}
    \begin{itemize}
      \item There is a finite $U \subseteq \reals^d$ such that, for all $p$, $U$ contains an optimal prediction.
      \item These $U$ partition $\simplex$ into finitely many polytopes.
    \end{itemize}
    \item Regret is linear on each polytope.
    \item Therefore $\Reg_L(u ; p)$ is a convex combination over the corners.
    \item So we can take $C = \max_{\text{corners $q$}} \alpha_q$.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{~}{}
  \begin{block}{Theorem 2}
    Let $L$ be a locally strongly convex surrogate with locally Lipschitz gradient.
    Suppose $(L,\psi)$ is consistent for $\ell$ with:
    \[  \Reg_{\ell}(\psi \circ h) \leq \zeta\Big( \Reg_L(h) \Big) . \]
   Then there exists $c > 0$ such that, for all small enough $\epsilon > 0$,
    \[ \zeta(\epsilon) \geq c \sqrt{\epsilon} . \]
  \end{block}

  \pause
  \emph{Proof idea:}
  \begin{itemize}
    \item Fix a ``boundary'' prediction $u_0$.
    \item Consider a conditional distributions $\{p_{\lambda} : 0 \leq \lambda \leq 1\}$
    \item \redem{Target regret} shrinks linearly as $\lambda \to 0$, but \blueem{surrogate regret} shrinks with $\sqrt{\lambda}$.
  \end{itemize}

  \pause
  Examples: exponential loss, Huber loss \hintright{via strengthenings}
\end{frame}
